---
title: "Assignment 2 - Adv_Machine_Learning"
author: "Mohamed Megahed"
date: "3/1/2020"
output: html_document
---

There are two ways to obtain word embeddings:

## Learning word embeddings with an embedding layer

```{r}
imdb_dir <- "C:/E/Advanced Machine Learning/Deep Learning book codes/deep-learning-with-r-notebooks-master/notebooks/aclImdb_v1/aclImdb"
train_dir <- file.path(imdb_dir, "train")

labels <- c()
texts <- c()

for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(train_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}
```

```{r}
library(keras)
maxlen <- 150                 # We will cut reviews after 150 words
training_samples <- 100       # We will be training on 100 samples
validation_samples <- 10000   # We will be validating on 10000 samples
max_words <- 10000            # We will only consider the top 10,000 words in the dataset

tokenizer <- text_tokenizer(num_words = max_words) %>% 
  fit_text_tokenizer(texts)

sequences <- texts_to_sequences(tokenizer, texts)

word_index = tokenizer$word_index
cat("Found", length(word_index), "unique tokens.\n")

data <- pad_sequences(sequences, maxlen = maxlen)

labels <- as.array(labels)
cat("Shape of data tensor:", dim(data), "\n")
cat('Shape of label tensor:', dim(labels), "\n")

# Split the data into a training set and a validation set
# But first, shuffle the data, since we started from data
# where sample are ordered (all negative first, then all positive).
indices <- sample(1:nrow(data))
training_indices <- indices[1:training_samples]
validation_indices <- indices[(training_samples + 1): 
                              (training_samples + validation_samples)]

x_train <- data[training_indices,]
y_train <- labels[training_indices]

x_val <- data[validation_indices,]
y_val <- labels[validation_indices]
```

```{r, echo=TRUE, results='hide'}
model <- keras_model_sequential() %>% 
  # We specify the maximum input length to our Embedding layer
  # so we can later flatten the embedded inputs
  layer_embedding(input_dim = 10000, output_dim = 8, 
                  input_length = maxlen) %>% 
  # We flatten the 3D tensor of embeddings 
  # into a 2D tensor of shape `(samples, maxlen * 8)`
  layer_flatten() %>% 
  # We add the classifier on top
  layer_dense(units = 1, activation = "sigmoid") 

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)

history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)
```

```{r}
plot(history)
```


## Using pre-trained word embeddings

```{r}
glove_dir = 'C:/E/Advanced Machine Learning/Deep Learning book codes/deep-learning-with-r-notebooks-master/notebooks/glove.6B'
lines <- readLines(file.path(glove_dir, "glove.6B.100d.txt"))

embeddings_index <- new.env(hash = TRUE, parent = emptyenv())
for (i in 1:length(lines)) {
  line <- lines[[i]]
  values <- strsplit(line, " ")[[1]]
  word <- values[[1]]
  embeddings_index[[word]] <- as.double(values[-1])
}

cat("Found", length(embeddings_index), "word vectors.\n")
```
Next, we'll build an embedding matrix that we can load into an embedding layer. It must be a matrix of shape `(max_words, embedding_dim)`, where each entry _i_ contains the `embedding_dim`-dimensional vector for the word of index _i_ in the reference word index (built during tokenization).
```{r}
embedding_dim <- 100

embedding_matrix <- array(0, c(max_words, embedding_dim))

for (word in names(word_index)) {
  index <- word_index[[word]]
  if (index < max_words) {
    embedding_vector <- embeddings_index[[word]]
    if (!is.null(embedding_vector))
      # Words not found in the embedding index will be all zeros.
      embedding_matrix[index+1,] <- embedding_vector
  }
}
```

Additionally, we'll freeze the weights of the embedding layer
```{r, echo=TRUE, results='hide'}
model1 <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_words, output_dim = embedding_dim, 
                  input_length = maxlen) %>% 
  layer_flatten() %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
```

### Load the GloVe embeddings in the model
```{r}
get_layer(model1, index = 1) %>% 
  set_weights(list(embedding_matrix)) %>% 
  freeze_weights()
```

Let's compile our model and train it:
```{r, echo=TRUE, results='hide'}
model1 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)

history1 <- model1 %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 32,
  validation_data = list(x_val, y_val)
)

save_model_weights_hdf5(model1, "pre_trained_glove_model.h5")
```

Let's plot its performance over time:
```{r}
plot(history1)
```

Finally, let's evaluate the model on the test data. First, we will need to tokenize the test data:
```{r}
test_dir <- file.path(imdb_dir, "test")

labels <- c()
texts <- c()

for (label_type in c("neg", "pos")) {
  label <- switch(label_type, neg = 0, pos = 1)
  dir_name <- file.path(test_dir, label_type)
  for (fname in list.files(dir_name, pattern = glob2rx("*.txt"), 
                           full.names = TRUE)) {
    texts <- c(texts, readChar(fname, file.info(fname)$size))
    labels <- c(labels, label)
  }
}

sequences <- texts_to_sequences(tokenizer, texts)
x_test <- pad_sequences(sequences, maxlen = maxlen)
y_test <- as.array(labels)
```

And let's load and evaluate the first model:

```{r}
model1 %>% 
  load_model_weights_hdf5("pre_trained_glove_model.h5") %>% 
  evaluate(x_test, y_test, verbose = 0)
```

We get an appalling test accuracy of 56%. Working with just a handful of training samples is hard!